The Milky Way Project: Leveraging Citizen Science and Machine Learning to Detect Interstellar Bubbles

### 2014-06-09 paper/ms.tex
Manuscript update

\newcommand{\todo}[1]{{\color[rgb]{0, .5, .5} TODO: #1}}
\newcommand{\um}[0]{$\mu$m}
\newcommand{\hii}[0]{H$_{\rm II}$}

@@

\begin{abstract}
We present Brut, an algorithm to identify bubbles in infrared images of the Galactic midplane. Brut is based on the Random Forest algorithm, and uses bubbles identified by $>35,000$ citizen scientists from the Milky Way Project to discover the identifying characteristics of bubbles in images from the \textit{Spitzer Space Telescope}. We demonstrate that Brut's ability to identify bubbles is comparable to expert astronomers. We use Brut to re-assess the bubbles in the Milky Way Project catalog, and find that $10-30\%$ of the objects in this catalog are non-bubble interlopers. Relative to these interlopers, high-reliability bubbles are more confined to the mid plane, and display a stronger excess of Young Stellar Objects along and within bubble rims. Furthermore, Brut is able to discover bubbles missed by previous searches -- particularly bubbles near bright sources which have low contrast relative to their surroundings. Brut demonstrates the synergies that exist between citizen scientists, professional scientists, and machine learning techniques. In cases where ``untrained'' citizens can identify patterns that machines cannot detect without training, machine learning algorithms like Brut can use the \emph{output} of citizen science projects as \emph{input} training sets, offering tremendous opportunities to speed the pace of scientific discovery.
\end{abstract}

@@

Unfortunately, due to their complex morphologies, bubbles -- like many features of the interstellar medium -- are difficult to identify and analyze. Existing catalogs of spatially extended bubbles have typically been identified visually \citep{Churchwell06, Churchwell07, Simpson12, Helfand06}. This has two main disadvantages. First, it is cumbersome and increasingly infeasible as datasets grow ever larger. Second, manual classification is inherently subjective and non-repeatable; humans are susceptible to fatigue, boredom, and subtle selection biases whose impact on the resulting catalog is difficult to calibrate. The problems associated manual bubble detection are germane to many analyses with a subjective component.

Machine learning techniques represent a promising solution to these problems. These techniques aim to construct computational models that can distinguish between different classes of objects, without domain-specific knowledge of what such objects represent -- in other words, they identify purely statistical differences between different populations of data. While such models are not typically useful as \textit{scientific} models, they can be very effective as \textit{computational} models to perform tasks like classification.

Our goal in this work is to apply machine learning techniques to the task of bubble detection, and to evaluate the potential of this approach. Using a catalog of known bubbles identified by the citizen scientists of the Milky Way Project \citep{Simpson12}, we ``teach'' an algorithm to identify bubbles in image data from the \textit{Spitzer} space telescope. We describe the design of this algorithm, which we call Brut, in Section \ref{sec:method}. In Section \ref{sec:expert}, we use a set of expert classifications to measure Brut's performance at bubble detection. In Section \ref{sec:prob}, we demonstrate that this detector produces useful probabilistic estimates for whether any particular image contains a bubble -- these probabilities correlate well with how astronomers classify similar regions. We use this detector to look for biases and incompleteness in existing bubble catalogs. This analysis yields a new catalog of high-probability bubbles, and we explore how the ensemble properties of this catalog differ from the old catalog. In Section \ref{sec:blind}, we apply Brut to the task of discovering bubbles missing from current catalogs. In Section \ref{sec:next_steps}, we consider how this approach applies more generally to future data analysis problems in astronomy.

@@

Generally speaking, a bubble is a shell-like, 1-30 parsec-scale cavity in the ISM, cleared by a combination of thermal overpressure, radiation pressure, and stellar winds. The basic structure of a bubble is shown in Figure \ref{fig:schematic}. \cite{Stromgren39} first derived the ionization structure around OBA stars in the ISM. The act of ionizing hydrogen heats gas from 100 to 10$^4$K, creating a strong overpressure that drives expansion into the surrounding medium. This expansion creates an overdense shell of gas along the expansion front. \cite{Castor75} and \cite{Weaver77} extended this analysis to include the effects of strong stellar winds from O and early B stars. The main effect of a stellar wind is to create a shocked, high-temperature (10$^6$K) region within the 10$^4$K ionization region. This shock sweeps up additional material within the ionization region, potentially creating a second shell. Though the term ``bubble'' originally referred to cavities cleared by stellar winds, modern observational studies of ``bubbles'' tend to include both these objects and classical \hii\, regions. Furthermore, \cite{Beaumont10} demonstrated that many bubbles are embedded in relatively thin clouds, and more closely resemble rings than spheres. Throughout this paper, we use the term ``bubble'' broadly to refer to any ring- or shell-like cavity cleared by a young or main sequence star.

@@

\cite{Churchwell06} and \cite{Churchwell07} carried out the first search for Bubbles in \emph{Spitzer} images of the Galactic plane, yielding  a catalog of some 600 bubbles located at $|\ell| < 65^\circ$, $|b| < 1^\circ$. These objects were identified by four astronomers manually searching through 3.5-8.0 \um\, images (they did not have access to 24 \um\, images). \cite{Churchwell06} noted that each astronomer possessed different selection biases, and cautioned that their catalog was likely incomplete.

In an attempt to overcome the inherent bias in manual classification, the web-based Milky Way Project \cite{Simpson12} enlisted over 35,000 citizen scientists to search for bubbles in \emph{Spitzer} data. The Milky Way Project (hereafter MWP) presented color-composite \emph{Spitzer} images at 4.5\um,  8\um, and 24\um, and asked citizen scientists to draw ellipses around potential bubbles. Figure \ref{fig:mwp} shows two typical images presented to the public. Each image in the Milky Way Project was 800x400 pixels, with a pixel scale ranging between 1.35$''$ and 6.75$''$ per pixel. The citizen science effort produced a dramatically larger catalog of $\sim 5000$ objects, nearly 10 times the number in the catalog of $\sim 600$ shells cataloged by the four astronomers of the \cite{Churchwell06, Churchwell07} surveys. The organizers of the MWP attribute this large increase to the 10,000-fold increase in human classifiers and the use of 24\um\, data to further emphasize bubble interiors. They estimate the MWP catalog to be 95\% complete, based on the falloff rate at which new objects were discovered. Still, they cautioned that this catalog is heterogenous, and probably affected by hard-to-measure selection biases.

@@

First, human classification is time consuming, and people-hours are a limited resource. Even by enlisting large communities of citizen scientists, data from next generation surveys will be too large to search exhaustively. For example, the $>$ 35,000 citizen scientists of the Milky Way Project classified roughly 45 GB of image data from \emph{Spitzer}. Current and next-generation astronomical datasets are many thousands of times larger than this, suggesting tens of millions of citizen scientists would be needed for similar exhaustive searches through tera- and petabyte datasets.

Second, many scientifically important tasks are not suitable for enlisting the public. Part of the appeal of the Milky Way Project is due to the fact that the \emph{Spitzer} images are beautiful, contain many bubbles, and are compelling to browse through. Searches for very rare objects, or tasks where the scientific justification is less apparent to a citizen scientist, may be less likely to entice large volunteer communities. \cite{Raddick13} considers the motivations of citizen scientists in greater detail.

@@

The downside to automated classification for tasks like bubble detection is that the algorithms typically require a training set of example objects to classify. Fortunately, the Milky Way Project catalog provides just such a training set. We build an automatic bubble detector from these data in the present paper. There are several reasons to do this:

@@

Our goal is to use the set of citizen-scientist-identified bubbles to build an automatic detector that, when presented with a region of the sky in the  \textit{Spitzer} {\sc glimpse} and {\sc mipsgal} surveys, accurately determines whether or not the image contains a bubble. Our approach, which we name Brut, is an example of a supervised learning problem. Here is a brief overview of the task:

@@

\subsection{Training Data}
\label{sec:method_training_data}
The MWP catalog is a natural choice for a training dataset. However, as shown in Figure \ref{fig:gallery}, the catalog as a whole contains many examples of objects that are not bubbles, and can compromise the training process. Thus, we manually curated a list of 468 objects in the Milky Way Project catalog which were clear examples of bubbles. We focus our attention on objects in the large bubble catalog, which contains 3744 out of 5106 total bubbles, with simimajor axes  $a \geq 13.5''$  \citep{Simpson12}. We focus on the large bubble catalog since the web interface used to build the small catalog was different, and may have different selection biases. The largest bubble in the MWP catalog has a semi major axis $a=10.3'$. This upper limit is set by the widest-field images presented to citizen scientists -- we discuss an implication of this cutoff in Section \ref{sec:prob}.

@@

A better set of negative examples includes more fields containing structure (Figure \ref{fig:bootstrap_neg}). We built such a collection in a bootstrapped fashion. Starting with random negative examples, we trained a classifier and used it to scan a large number (20,000) of negative fields. We then discarded half of the initial negative examples (those classified most confidently as not containing bubbles), and replaced them with a random sample of the mis-classified examples. We repeated this process several times, but found that one iteration was usually sufficient to build a good set of training data.

@@

\subsection{Feature Extraction}
\label{sec:method_feature_extraction}
Supervised classification algorithms distinguish between classes of objects based on numerical \textit{feature vectors} that summarize each object. Feature vector design is often the most important stage of supervised classification and, unfortunately, also the hardest. The ideal feature vector captures the differences between each class of objects, so that examples from different classes occupy different sub-regions of feature space.

In the case of image classification, the most obvious choice for a feature vector is simply the numerical value of the pixels themselves. This turns out to be a poor choice, because extended objects like bubbles are characterized by correlations between hundreds or thousands of pixels, and any individual pixel is virtually irrelevant to the classification task. Machine learning algorithms generally perform better when individual elements of the feature vector are more important to the classification, and less dependent on the value of other feature vector elements.

Our feature vectors have been designed based on insights in the automated face detection literature \citep{ViolaJones}.  The basic strategy is to encode a very large number ($\sim$40,000) of generic statistics about extended structures at various positions and scales. This strategy seems to be more effective than trying to tune a feature vector to the specific object being identified. While most elements in the feature vector will have no predictive power in the classification, the appropriate learning algorithm will simply ignore this information.

To construct our feature vector for a given region in the sky, we first extract a (40x40) pixel postage stamp of the field from IRAC 8 \um\, and MIPS 24 \um\, images. Following a scheme similar to \cite{Simpson12}, these images are intensity clipped at the 1 and 97th percentile, normalized to maximum intensity of 1, and passed through a square root transfer function. This scaling tends to do a good job of emphasizing ISM structure, making bubbles more visible to the eye. Then, the following properties were extracted from each postage stamp, and concatenated to form the feature vector:

\begin{enumerate}
\item The wavelet coefficients from the Discrete Cosine and Daub4 wavelet transforms. These coefficients encode how much power is in various spatial frequencies.
\item The image dot product\footnote{i.e., the sum of the pixel-by-pixel product of two images} of each image with 49 template images of circular rings of different size and thickness. Bubbles are morphologically similar to these templates, and presumably have larger dot products than a random intensity distribution.
\item The byte size of each compressed postage stamp image -- images substantially more or less smooth than a bubble compress to smaller and larger files, respectively.
\item The DAISY features of the image \citep{DAISY}. DAISY features are derived from measurements of local intensity gradients, and encode information about simple shapes in images; they are conceptually similar to edge detection filters.
\end{enumerate}

\subsection{Training}
While there are many algorithms designed to learn classifications form a training set of feature vectors, we focus here on the Random Forest algorithm \citep{Breiman01}. Random Forests are aggregates of a simpler learning algorithm called a decision tree. A decision tree is conceptually similar to the game twenty questions: in this game, one player asks a series of yes/no questions in an attempt to deduce the identity of a hidden object selected by another player. Likewise, a decision tree classifies an object by performing several boolean tests on elements of that object's feature vector (e.g., $F_3 > 5$). These tests are arranged in a binary tree. The decision tree algorithm traverses this tree for each object, traversing the left or right subtree depending on whether each boolean test is True or False. The leaves of the tree contain the final classification (in the present context, ``bubble'' or ``not bubble'').

The decision tree is constructed from the training data, where the feature vectors and classification are both known. Trees are built one node at a time, in a greedy fashion -- that is, at every step of the learning process, a node is added to the tree to maximize the tree's classification performance on the training data. Several different heuristics can be used to measure classification performance.

>>>>

\newcommand{\todo}[1]{{\color[rgb]{0, .5, .5} TODO: #1}}
\newcommand{\um}[0]{$\mu$m}
\newcommand{\hii}[0]{{\sc H\,ii}}
\newcommand{\hi}[0]{{\sc H\,i}}

@@

\begin{abstract}
We present Brut, an algorithm to identify bubbles in infrared images of the Galactic midplane. Brut is based on the Random Forest algorithm, and uses bubbles identified by $>35,000$ citizen scientists from the Milky Way Project to discover the identifying characteristics of bubbles in images from the \textit{Spitzer Space Telescope}. We demonstrate that Brut's ability to identify bubbles is comparable to expert astronomers. We use Brut to re-assess the bubbles in the Milky Way Project catalog, and find that $10-30\%$ of the objects in this catalog are non-bubble interlopers. Relative to these interlopers, high-reliability bubbles are more confined to the mid plane, and display a stronger excess of Young Stellar Objects along and within bubble rims. Furthermore, Brut is able to discover bubbles missed by previous searches -- particularly bubbles near bright sources which have low contrast relative to their surroundings. Brut demonstrates the synergies that exist between citizen scientists, professional scientists, and machine learning techniques. In cases where ``untrained'' citizens can identify patterns that machines cannot detect without training, machine learning algorithms like Brut can use the \emph{output} of citizen science projects as \emph{input} training sets, offering tremendous opportunities to speed the pace of scientific discovery. A hybrid model of machine learning combined with crowdsourced training data from citizen scientists can not only classify large quantities of data, but also address the weakness of each approach if deployed alone.
\end{abstract}

@@

Unfortunately, due to their complex morphologies, bubbles -- like many features of the interstellar medium -- are difficult to identify and analyze. Existing catalogs of spatially extended bubbles have typically been identified visually \citep{Hu81, Marston96,  Kiss04, Helfand06, Churchwell06, Churchwell07, Konyves07, Simpson12}. This has two main disadvantages. First, it is cumbersome and increasingly infeasible as datasets grow ever larger. Second, manual classification is inherently subjective and non-repeatable; humans are susceptible to fatigue, boredom, and subtle selection biases whose impact on the resulting catalog is difficult to calibrate. The problems associated with manual bubble detection are germane to many analyses with a subjective component.

Machine learning techniques represent a promising solution to these problems. These techniques aim to construct models that can distinguish between different classes of objects, without domain-specific knowledge of what such objects represent -- in other words, they identify purely statistical differences between different populations of data. While such models are not typically useful as \textit{scientific} models, they can be very effective as \textit{computational} models to perform tasks like classification.

Our goal in this work is to apply machine learning techniques to the task of bubble detection, and to evaluate the potential of this approach. Using a catalog of known bubbles identified by the citizen scientists of the Milky Way Project \citep{Simpson12}, we ``teach'' an algorithm to identify bubbles in image data from the \textit{Spitzer Space Telescope}. We describe the design of this algorithm, which we call Brut, in Section \ref{sec:method}. In Section \ref{sec:expert}, we use a set of expert classifications to measure Brut's performance at bubble detection. In Section \ref{sec:prob}, we demonstrate that this detector produces useful probabilistic estimates for whether any particular image contains a bubble -- these probabilities correlate well with how astronomers classify similar regions. We use this detector to look for biases and incompleteness in existing bubble catalogs. This analysis yields a new catalog of high-probability bubbles, and we explore how the ensemble properties of this catalog differ from the old catalog. In Section \ref{sec:blind}, we apply Brut to the task of discovering bubbles missing from current catalogs. In Section \ref{sec:next_steps}, we consider how this approach applies more generally to future data analysis problems in astronomy.

@@

Generally speaking, a bubble is a shell-like, 1-30 parsec-scale cavity in the ISM, cleared by a combination of thermal overpressure, radiation pressure, and stellar winds. The basic structure of a bubble is shown in Figure \ref{fig:schematic}. \cite{Stromgren39} first derived the ionization structure around OBA stars in the ISM. Stellar radiation both ionizes and heats gas up to $\sim$ 10$^4$K, creating a strong overpressure that drives expansion into the surrounding medium. This expansion creates an overdense shell of gas along the expansion front. \cite{Castor75} and \cite{Weaver77} extended this analysis to include the effects of strong stellar winds from O and early B stars. The main effect of a stellar wind is to create a shocked, high-temperature (10$^6$K) region within the 10$^4$K ionization region. This shock sweeps up additional material within the ionization region, potentially creating a second shell. Though the term ``bubble'' originally referred to cavities cleared by stellar winds, modern observational studies of ``bubbles'' tend to include both these objects and classical \hii\, regions. Furthermore, \cite{Beaumont10} demonstrated that many bubbles are embedded in relatively thin clouds, and more closely resemble rings than spheres. Throughout this paper, we use the term ``bubble'' broadly to refer to any ring- or shell-like cavity cleared by a young or main sequence star.

@@

\cite{Churchwell06} and \cite{Churchwell07} carried out the first search for bubbles in \emph{Spitzer} images of the Galactic plane, yielding  a catalog of some 600 bubbles located at $|\ell| < 65^\circ$, $|b| < 1^\circ$. These objects were identified by four astronomers manually searching through 3.5-8.0 \um\, images (they did not have access to 24 \um\, images). \cite{Churchwell06} noted that each astronomer possessed different selection biases, and cautioned that their catalog was likely incomplete.

In an attempt to overcome the inherent bias in manual classification, the web-based Milky Way Project \citep{Simpson12} enlisted over 35,000 citizen scientists to search for bubbles in \emph{Spitzer} data. The Milky Way Project (hereafter MWP) presented color-composite \emph{Spitzer} images at 4.5\um,  8\um, and 24\um, and asked citizen scientists to draw ellipses around potential bubbles. Figure \ref{fig:mwp} shows two typical images presented to the public. Each image in the MWP was 800x400 pixels, with a pixel scale ranging between 1.35$''$ and 6.75$''$ per pixel. The citizen science effort produced a dramatically larger catalog of $\sim 5000$ objects, nearly 10 times the number in the catalog of $\sim 600$ shells cataloged by the four astronomers of the \cite{Churchwell06, Churchwell07} surveys. The organizers of the MWP attribute this large increase to the 10,000-fold increase in human classifiers and the use of 24\um\, data to further emphasize bubble interiors. They estimate the MWP catalog to be 95\% complete, based on the falloff rate at which new objects were discovered. Still, they cautioned that this catalog is heterogenous, and probably affected by hard-to-measure selection biases.

@@

First, human classification is time consuming, and people-hours are a limited resource. Even by enlisting large communities of citizen scientists, data from next generation surveys will be too large to search exhaustively. For example, the $>$ 35,000 citizen scientists of the MWP classified roughly 45 GB of image data from \emph{Spitzer}. Current and next-generation astronomical datasets are many thousands of times larger than this, suggesting tens of millions of citizen scientists would be needed for similar exhaustive searches through tera- and petabyte datasets.

Second, many scientifically important tasks are not suitable for enlisting the public. Part of the appeal of the MWP is due to the fact that the \emph{Spitzer} images are beautiful, contain many bubbles, and are compelling to browse through. Searches for very rare objects, or tasks where the scientific justification is less apparent to a citizen scientist, may be less likely to entice large volunteer communities. \cite{Raddick13} considers the motivations of citizen scientists in greater detail.

@@

There have been a few previous attempts to detect shell-like structures in the ISM using automated techniques. One approach involves correlating observational data with templates, to look for characteristic morphologies. \cite{Thilker98}, for example, used position-position-velocity templates of expanding shells to search for bubbles in \hi\, data. This technique is also used by \cite{Mashchenko95, Mashchenko02}. \cite{Ehlerova05, Ehlerova13} pursued a more generic approach -- they looked for the cavities interior to \hi\, shells by identifying extended regions of low emission. \cite{Daigle03, Daigle07} searched for \hi\, shells by training a neural network to classify spectra as containing outflow signatures, based on a training set of spectra from known shells. They pass spectral cubes through this classifier, and look for connected regions of outflow signatures.

It is difficult to robustly classify ISM structures using templates, due to the heterogeneity and irregularity of the ISM -- simple shapes like expanding spheres are often poor approximations to the true ISM. The method pursued by Ehlerov{\'a} et al. is more robust in this respect but, as \cite{Daigle07} noted, is only sensitive to fully-enclosed bubbles. Real bubbles are often broken or arc-like. In this respect the neural network approach that Daigle et al. pursue is an improvement, since neural networks make fewer hard-coded assumptions about the defining characteristics of a bubble, and instead use the data to discover discriminating features. There are a few drawbacks to this approach. First, Daigle et al.'s neural network classifies a single spectrum at a time, with no complementary information about the spatial structure of emission on the sky. Second, the neural networks in \cite{Daigle07} are trained on data from 11 bubbles from the Canadian Galactic Plane Survey, and thus potentially biased towards the peculiarities of this small sample. Finally, neural networks have a large number of tunable parameters related to the topology of the network, and the optimal topology has a reputation for being difficult to optimize.

Compared to these efforts, the approach we pursue here is similar to the work of \cite{Daigle07}. However, our methodology differs in a few key respects. First, the MWP catalog provides a two orders-of-magnitude larger sample of known bubbles with which to train an automatic classifier. Second, we use the more recent random forest algorithm \citep{Breiman01}, which has a smaller number of tunable parameters that are easier to optimize. Finally, we focus on detecting morphological signatures of shells in \emph{Spitzer} data -- infrared bubbles are visually more prominent than \hi\, shells, and thus potentially easier to classify.

There are several reasons to do build a bubble classifier based on MWP data:

@@

Our goal is to use the set of citizen-scientist-identified bubbles to build an automatic detector that, when presented with a region of the sky in the  \textit{Spitzer} {\sc glimpse} and {\sc mipsgal} surveys, accurately determines whether or not the image contains a bubble. Our approach, which we name Brut\footnote{\url{http://github.com/ChrisBeaumont/brut}}, is an example of a supervised learning problem. Here is a brief overview of the task:

@@

\subsection{Random Forest Classification}
Brut uses the Random Forest classification algorithm \citep{Breiman01} to discriminate between images of bubbles and non-bubbles. Random Forests are aggregates of a simpler learning algorithm called a decision tree. A decision tree is a data structure which classifies feature vectors by computing a series of constraints, and propagating vectors down the tree based on whether these constraints are satisfied. For example, Figure \ref{fig:decision_schematic} shows a simple decision tree for classifying animals as cats or not. Here, the feature vectors are four properties of each animal observed: number of legs, tails, height, and mass. Two example feature vectors are shown, and propagate to the outlined classification nodes. In the context of Brut, each classification object is a square region in the {\sc glimpse} survey. Each feature vector is  a set of morphological properties extracted from one such region, and the decision tree predicts whether each region is filled by a bubble.

\begin{figure}[h!]
\includegraphics[width=3in]{decision_schematic}
\caption{Schematic example of a decision tree, to classify observations of animals as cats or not. Two feature vectors are shown at the top, and their classifications are shown as the red and blue outlined nodes of the decision tree.}
\label{fig:decision_schematic}
\end{figure}

Decision trees are constructed using an input training set of pre-classified feature vectors. During tree construction, a quality heuristic is used to rate the tree. A few heuristics are common, which consider both the classification accuracy and the complexity of the tree itself -- highly complex trees are more prone to over fitting, and thus disfavored \citep{Ivezic14}. We treat the choice of specific heuristic as a hyper-parameter, which we discuss below. Decision trees are constructed one node at a time, in a ``greedy'' fashion. That is, at each step in the learning process, a new boolean constraint is added to the tree, to maximally increase the score of the quality heuristic. This process repeats until the quality heuristic reaches a local maximum.

On their own, decision trees are prone to over-fitting the training data by adding too many nodes. In a more traditional fitting context, this is analogous to adding too many terms to a polynomial fit of noisy data -- in both situations, the model describes the input data very well, but does not generalize to new predictions. Random Forests were designed to overcome this problem \citep{Breiman01}. Random forests are ensembles of many decision trees, built using different random subsets of the training data. The final classification of a Random Forest is simply the majority classification of the individual trees. Since over-fitting is sensitive to the exact input data used, each individual tree in a Random Forest over-fits in a different way. By averaging the classifications of each tree, the Random Forest mitigates over-fitting. Random Forests have proven effective in many machine learning contexts \citep{Kuhn13}.

In addition to good performance and robustness against over fitting, the Random Forest algorithm has several other advantageous properties: it is possible to interpret how important each element in the feature vector is to the classification (Section \ref{sec:feature_importance}), the algorithm naturally ignores irrelevant elements of the feature vector, and it is conceptually and computationally simple.

It is worthwhile to compare the properties of learning algorithms like the Random Forest classifier to more traditional, ``parametric'' models commonly used in astronomy. Parametric models are often constructed by hand, and derived from an underlying astrophysical model -- the fit to the power spectrum of the Cosmic Microwave Background is a quintessential example of such a model. The structure of a Random Forest model, on the other hand, is set by the clustering properties of the data themselves. The main advantage of the Random Forest approach is that the complexity of the model is to a large extent driven by the complexity of the training data. That is, Random Forests can model complex morphological differences between bubble and non-bubble images, which would otherwise be difficult to express in a manually-constructed model of bubble structure. This is a crucial property when considering the irregularity of the ISM.

\subsection{Feature Extraction}
\label{sec:method_feature_extraction}

The individual ``objects'' that Brut classifies are square regions of the sky, and the goal of the classifier is to determine whether each region is filled by a bubble. Each field is identified by 3 numbers: the latitude and longitude of the center of the field, and the size of the field. We decompose the {\sc glimpse} survey coverage into fields of 18 different sizes, logarithmically spaced from 0.02$^\circ$ to 0.95$^\circ$. At each size scale, we define and classify an overlapping grid of fields. Neighboring tiles in this grid are separated by $1/5$ of the tile size.

As a preprocessing step, we extracted two-color postage stamps for each field (at 8 \um\, and 24 \um), and resampled these postage stamps to (40x40) pixels. Following a scheme similar to \cite{Simpson12}, these images were intensity clipped at the 1 and 97th percentile, normalized to maximum intensity of 1, and passed through a square root transfer function. The intensity scaling tends to do a good job of emphasizing ISM structure, making bubbles more visible to the eye. Likewise, the (40x40) pixel resolution was chosen because it is reasonably small, yet has enough resolution that postage stamps of known bubbles are still recognizable as such by humans.  Figure \ref{fig:preprocess} shows four preprocessed fields towards known bubbles\footnote{other figures in this paper use the same intensity scaling scheme, but use a (200x200) pixel resolution for presentation purposes}. The goal of preprocessing is to standardize the appearance of bubbles as much as possible, across different size scales and ambient intensities. All subsequent stages of Brut work exclusively with these images, as opposed to the unscaled pixel data.

\begin{figure}[h!]
\includegraphics{preprocess}
\caption{Examples of preprocessed images for fields towards four known bubbles.}
\label{fig:preprocess}
\end{figure}

The input to a Random Forest classifier is a numerical feature vector that describes the properties of each region. The ideal feature vector captures the differences between each class of objects, so that objects from different classes occupy different sub-regions of feature space. Designing an effective feature vector is often the most important determinant of classification performance.

The most obvious choice for a feature vector is simply the numerical value of the pixels in each preprocessed postage stamp. This turns out to be a poor choice, because extended objects like bubbles are characterized by correlations between hundreds or thousands of pixels, and any individual pixel is virtually irrelevant to the classification task. Machine learning algorithms generally perform better when individual elements of the feature vector are more important to the classification, and less dependent on the value of other feature vector elements.

Our feature vectors have been designed based on insights in the automated face detection literature \citep{ViolaJones}.  The basic strategy is to encode a very large number ($\sim$40,000) of generic statistics about image structures at various positions and scales. This strategy seems to be more effective than trying to tune a feature vector to the specific object being identified. While most elements in the feature vector will have no predictive power in the classification, the Random Forest classifier is capable of ignoring this information, and finding the elements in the feature vector with the most predictive power.

The following quantities are extracted from each postage stamp, and concatenated to form a feature vector.

\begin{enumerate}
\item The wavelet coefficients from the Discrete Cosine and Daub4 wavelet transforms \citep{Press07}. These coefficients represent the weights used in representing an image as a linear combination of basis functions. The basis functions in the Discrete Cosine transform are cosine functions, and thus encode information about power at different spatial frequencies. By contrast, the Daub4 wavelets are spatially compact, and thus better encode spatially-isolated structures in an image. These wavelet-derived features resemble the features employed for face detection \citep{ViolaJones}.
\item The image dot product\footnote{i.e., the sum of the pixel-by-pixel product of two images} of each image with 49 template images of circular rings of different size and thickness. Bubbles are morphologically similar to these templates, and tend to have larger dot products than a random intensity distribution. All ring templates are centered within the postage stamp. They span seven radius values ranging from 1-20 pixels, and seven thicknesses (2, 4, 6, 8, 10, 15, and 20 pixels).
\item The byte size of each compressed postage stamp image -- images substantially more or less smooth than a bubble compress to smaller and larger files, respectively.
\item The DAISY features of the image \citep{DAISY}. DAISY features are derived from measurements of local intensity gradients, and encode information about simple shapes in images; they are conceptually similar to edge detection filters.
\end{enumerate}

We arrived at these specific feature vectors through experimentation, because they empirically yield effective classifications. Feature vector design is often the most subjective component of many machine learning tasks, and other features may better model the differences between bubbles and non-bubbles.

\subsection{Edge-of-field limitations}
\label{sec:feature_edges}
We found that classification performance improved if, in addition to extracting feature vectors at a single $(\ell, b, d)$ field, we also extracted features at $(\ell, b, 2 \times d)$, $(\ell, b, d / 2)$ and $(\ell, b + d/ 2, d)$ -- here, $d$ is the bubble diameter. This ensemble of fields gives the classifier context about the region surrounding a bubble.

Thus, in our approach, Brut extracts feature vector information from a region of the sky twice as large as the field it is classifying. This imposes a latitutde-dependent limit on the largest field Brut is able to classify. Because the {\sc glimpse} coverage is limited to $|b| < 1^\circ$, fields where $|b| + d > 1^\circ$ partially fall outside the survey area. We cannot derive proper feature vectors for these regions, and thus bubbles like this are undetectable by Brut. We return to this fact in Section \ref{sec:prob}.

\subsection{Training Data}
\label{sec:method_training_data}

Building a Random Forest requires providing a set of pre-classified feature vectors. The MWP catalog is a natural choice for building such a training set. However, as shown in Figure \ref{fig:gallery}, the catalog as a whole contains many examples of objects that are not bubbles, and can compromise the training process. Thus, we manually curated a list of 468 objects in the MWP catalog which were clear examples of bubbles. We focus our attention on objects in the large bubble catalog, which contains 3744 out of 5106 total bubbles, with semimajor axes  $a \geq 13.5''$  \citep{Simpson12}. We focus on the large bubble catalog since the web interface used to build the small catalog was different, and may have different selection biases. The largest bubble in the MWP catalog has a semi major axis $a=10.3'$. This upper limit is set by the widest-field images presented to citizen scientists -- we discuss an implication of this cutoff in Section \ref{sec:prob}. The full training set of positive examples is obtained by extracting a field centered on each of these 468 clean bubbles, with a size equal to the major axis of each bubble.

@@

A better set of negative examples includes more fields containing structure (Figure \ref{fig:bootstrap_neg}). We built such a collection in a bootstrapped fashion. We began with a random set of negative fields, distributed uniformly in latitude and longitude, with sizes drawn from the size distribution of the positive training set. We trained a classifier with these examples and used it to scan 20,000 bubble-free regions. We then discarded half of the initial negative examples (those classified most confidently as not containing bubbles), and replaced them with a random sample of the mis-classified examples. We repeated this process several times, but found that one iteration was usually sufficient to build a good set of training data.

Our final training set consists of 468 examples of bubbles identified by the Milky Way Project, and 2289 examples of non-bubble fields. We found that an approximate five-fold excess of negative examples yielded a more discriminating classifier with lower false positive rates, compared to a training set with equal number of positive and negative examples. Likewise, using dramatically more negative examples compromised the classifier's ability to positively identify known bubbles. The number of positive examples was set by the number of clear bubbles in the Milky Way Project catalog itself (as judged by the lead author).

@@

\subsection{Partitioning the Classifier}
\label{sec:partition}
Instead of building a single Random Forest classifier, we trained three forests on different subsets of the sky. Each forest was trained using examples from 2/3 of the {\sc glimpse} survey area, and used to classify the remaining 1/3. The motivation for doing this is to minimize the chance of over-fitting, by ensuring that the regions of the sky used to train each classifier do not overlap the regions of sky used for final classification.

The training zones used to classify each forest are interleaved across all longitudes. For example, Figure \ref{fig:zone} depicts how the $0^\circ < \ell < 10^\circ$ region is partitioned -- the shaded regions denote the portion of the sky used to train each forest, and the white region shows the zone each forest is responsible for classifying.

\begin{figure}[h!]
\includegraphics{zones}
\caption{Illustration of the zones used to train each Random Forest. Each forest is trained using the data from the shaded regions in its zone, and used to classify the light regions.}
\label{fig:zone}
\end{figure}

\subsection{Hyperparameter Optimization}
The Random Forest algorithm has a small number of tunable parameters:

\begin{enumerate}
\item  The heuristic used to measure classification quality when building a decision tree. We considered three heuristics: the Gini impurity, information gain, and gain ratio criteria. All of these provide similar measurements of how well a Decision tree partitions a training dataset -- \citep{Ivezic14} discuss these criteria in greater detail.
\item The number of individual decision trees to include in the Random Forest. We considered forests with 200, 400, 800, and 1600 trees.
\item A stopping criterion, when building individual decision trees. If this criteria is set to one, new nodes will always be added to a tree, until the quality heuristic reaches a local maximum. If this number is set to a value $c > 1$, then it prevents a subtree from being added to a given node in the decision tree unless at least $c$ training examples propagate to that node. This criteria can suppress over-fitting in individual trees, though it seems to be less important in the context of Random Forests. We considered values of 1, 2, and 4.
\end{enumerate}

We explored the impact of these hyperparameter settings via cross validation. During cross validation, we split the training examples into two groups: a primary set with 154 bubble examples and 1462 non-bubble examples, and a validation set of 157 bubble and 10000 non-bubble examples. We trained a Random Forest with a particular choice of hyper parameters, and measured the accuracy and false-positive rate of the classifier on the validation set. We found that a forest using the Information Gain heuristic, 800 trees, and $c=4$ yielded the best performance. The quality heuristic and forest size are stronger determinants of classification performance than $c$. Classifier performance largely converged at 800 trees. While forests with 1600 trees did no worse than 800-tree forests, classification time was twice as slow. Once we converged on an optimal set of hyperparameters, we trained the final three Random Forest classifiers using the scheme discussed above.

\subsection{Feature Importance}
\label{sec:feature_importance}
The Random Forest algorithm measures the importance of each element in the feature vector during training. Intuitively, elements in the feature vector which appear in a large number of nodes (see Figure \ref{fig:decision_schematic}) are more important to the classification. Likewise, elements which appear closer to the root of the tree are more important, since a larger fraction of examples propagate through those nodes. More formally, feature importance is defined by the average improvement in the quality heuristic for all nodes associated with a feature, weighted by the number of training examples which propagate through those nodes. Feature importance is more meaningful in a relative sense than an absolute sense -- features with higher importance scores have a greater impact on the behavior of the decision tree.

Table \ref{tab:importance} summarizes the importance of each category of features used in Brut -- it shows the average and sum of the importances for all feature elements in each category. Both columns are normalized by the maximum importance value. The features derived from ring templates have the highest total importance -- this is further evidenced by the fact that 19 out of the 20 most important single features are ring features (the 15th most important feature is a wavelet feature). While the average importance of the compression feature is slightly greater, there are far fewer compression features than ring features.

\begin{deluxetable}{lrr}
\tablecolumns{3}
\tablewidth{0in}
\tabletypesize{\scriptsize}
\tablecaption{Feature importances}
\tablehead{\colhead{Feature Category} & \colhead{Mean Importance} & \colhead{Total Importance}}
\startdata
ring	&  0.945 & 	 1.000 \\
compression &  1.000	 &  0.014 \\
wavelet &  0.010 	&  0.292\\
daisy &	 0.007	&  0.085 \\
\enddata
\label{tab:importance}
\end{deluxetable}

\subsection{Using the classifier}
\label{sec:using_classifier}
